---
title: "Analyst Tasks"
output:
  html_document: default
  pdf_document: default
---

```{r}
# load packages
set.seed(9999)
library(tidyverse)
library(readxl)
library(mice)
library(scales)
library(ggrepel)
library(rstatix)
```
**Task 1.**
```{r}
test_thera <- read_excel('Research Analyst_16SEPT2021.xlsx')
```

**Task 2.**
```{r}
thera_variable_names <- colnames(test_thera)
thera_variable_names
```

**Task 3.**  
The number of rows is 1079, and the number of columns is 20.
```{r}
thera_nrow <- nrow(test_thera)
thera_ncol <- ncol(test_thera)
```

**Task 4.**
```{r}
tail(test_thera$age, n = 6)
```

**Task 5.**
```{r}
test_thera <- test_thera %>% rename(sex = gender, "blood pressure" = "systolic bp preoperative" )
thera_variable_names <- colnames(test_thera)
```

**Task 6.**
```{r}
test_thera$`income bracket` = recode(test_thera$`income bracket`,
                                     '1' = '<10,000',
                                     '2' = '10,000 to 20,000',
                                     '3' = '20,001 to 30,000',
                                     '4'='30,001 to 40,000',
                                     '5'='>40,000' )
```
**Task 7.**
```{r}
head(test_thera$`income bracket`, n= 6)
```
**Task 8.**
Seed was generated from no. 9999

Imputations closest to the mean of columns were selected: 58.14 for duration of operation and 137.7 for blood pressure, respectively.

Imputation 2 has the closest mean to duration of operation at 58.42, and the closest mean to blood pressure at 139.2, thus Imputation 2 was selected.
```{r}
summary(test_thera$`duration of operation`)
summary(test_thera$`blood pressure`)

thera_imputed <- select(test_thera, `duration of operation`, `blood pressure`)

# change space to underscore to run through MICE
thera_imputed <- thera_imputed %>% rename(duration_of_operation = "duration of operation",
                                          blood_pressure = "blood pressure")

#Using predictive mean matching method from mice. Generating 5 datasets
thera_imp <- mice(thera_imputed, m= 5, method = c("pmm","pmm"), maxit = 20)


summary(thera_imp$imp$duration_of_operation)
summary(thera_imp$imp$blood_pressure)

selected_imp <- complete(thera_imp,2)
summary(selected_imp)
```

**Task 9.**

The 5 chosen variables are discharge status, blood pressure, age, sex, type of surgery
```{r}

task9_variables <- test_thera %>% select(`blood pressure`, `discharge status`, age, sex, `type of surgery`)

task9_variables <- task9_variables %>% mutate(`discharge status` = as.factor(`discharge status`),
                                              sex = as.factor(sex),
                                              `type of surgery` = as.factor(`type of surgery`))

summary(task9_variables)

#Type of Surgery
TS_data<-  task9_variables %>% group_by(`type of surgery`)%>% summarise(count = n())

Typeofsurgery_graph<- ggplot(TS_data, aes(x="", y= count, fill=`type of surgery`)) +
geom_bar(stat= "identity",width = 1, color = "white",
size = 1)+ coord_polar("y", start=0) +
geom_text(aes(label = paste0(count," (",
          percent(count / sum(count)),
          ")")), position = position_stack(vjust = 0.5)) +
          ggtitle("Total Operations Completed by Type") +
          scale_fill_discrete(name = "Surgery Types Performed") + 
          theme(axis.title.y=element_blank(),  axis.ticks.y=element_blank())

Typeofsurgery_graph

#Age and Sex
AS_data <- task9_variables %>% group_by(age, sex) %>% summarise(count=n())

Age_distribution_by_sex <- ggplot(AS_data, aes(x=age, y=count, fill=sex)) +
                                  geom_bar(stat="identity") + 
                                  scale_x_continuous(breaks = pretty_breaks(n=9)) + 
                                  ggtitle("Age Distribution of MORG Trial Patients")
Age_distribution_by_sex

#Discharge Status
DIS_data <- task9_variables %>% group_by(`discharge status`) %>% summarise(count=n())

Discharge_ratio <- ggplot(DIS_data, aes(x="",y=count, fill=`discharge status`)) + 
                    geom_bar(stat="identity") + 
                    geom_text(aes(label=paste0(count)), position=position_stack(vjust=0.5)) +
                    ggtitle("Patient Status on Discharge") +
                    scale_fill_discrete(name="Patient Status") +
                    theme(plot.title=element_text(hjust = 0.5), 
                          axis.title.y=element_blank(),
                          axis.ticks.y=element_blank()) +
                    coord_flip()

Discharge_ratio

#Blood Pressure

BP_distribution <- ggplot(task9_variables, aes(x=`blood pressure`)) + 
                     geom_histogram(color="darkblue", fill="lightblue") + 
                     scale_x_continuous(breaks=pretty_breaks(n=9)) + 
                     ggtitle("Blood Pressure Distribution of MORG Trial Patients") + 
                     theme(plot.title=element_text(hjust=0.5)) + 
                     xlab("Blood Pressure (mmHg)")

BP_distribution
```

**Task 10.**

Tested for significant difference between high risk and low risk populations for all variables. To test the difference in means, a t-test was selected for numeric variables, while a chi- squared test was selected for categorical variables. Both tests can help visualize differences between both groups and provide a probability of the significance of the difference.

Significant difference was found between groups for blood pressure, GCS leak, Age, Duration of Operation, Sex, and Lumbar Drainage

```{r}
test_thera <- test_thera %>% mutate(risk_status = as.factor(case_when(`days in the NICU` <= 2 ~ "low risk",
                                                            `days in the NICU` > 2 ~ "high risk")))
#Performing multiple T-tests.
#T Tests are used for numeric variables
t1<- t_test(test_thera, `blood pressure` ~ risk_status )
t2<- t_test(test_thera, GCS ~ risk_status)
t3<- t_test(test_thera, age ~ risk_status)
t4<- t_test(test_thera, `duration of operation` ~ risk_status)

output_numeric_var <- bind_rows(t1,t2,t3,t4)
output_numeric_var

#Chi SQ Tests used for categorical variables
chisq.test(test_thera$risk_status,test_thera$`income bracket`)
chisq.test(test_thera$risk_status,test_thera$`diabetes mellitus`)
chisq.test(test_thera$risk_status,test_thera$`CSF leak`)
chisq.test(test_thera$risk_status,test_thera$`type of surgery`)
chisq.test(test_thera$risk_status,test_thera$sex)
chisq.test(test_thera$risk_status,test_thera$`discharge status`)
chisq.test(test_thera$risk_status,test_thera$`glucorticoids`)
chisq.test(test_thera$risk_status,test_thera$`SSI`)
chisq.test(test_thera$risk_status,test_thera$`lumbar drainage`)
chisq.test(test_thera$risk_status,test_thera$`entry of paranasal sinus`)
```

**Task 11.**

Tested univariate models between risk status and each variable in the dataset (excluding NICU time and patient ID)
Models significant at a 0.05 critical level were included in the multivariate model.
Any terms not significant in the multivariate model were excluded in the final multivariate model

Univariate Results:
As variables:
1. Type of Surgery, Use of Post-Operative Drain, Entry into paranasal sinus, diabetes mellitus, SSI, discharge status, glucorticoids, income bracket were not significant at a 0.05 critical level and will be excluded from the multivariate model

2. Age, sex,csf, duration of operation, GCS leak, lumbar drainage, blood pressure were significant at a 0.05 critical level and will be included in the multivariate model

For sake of cleanliness, variable interactions were not included in the multivariate model.

The variables: 'CSF leak' and GCS were not significant in the multivariate model and were thus excluded.

Final Model Results:
An increase in age, being a male, having lumbar drainage used, and an increase in blood pressure all contribute to a higher risk of being placed in the "high-risk" status. However, an increase in the duration of operation does not contribute to a higher risk of being placed in the "high-risk status" in the model.

```{r}

#Turned risk status into binary variable
test_thera$risk_status_binary <- ifelse(test_thera$risk_status == "high risk", 1, 0)

#Logarithmic Regression will be used as Risk Status is a Binary Variable

model_age <- glm(risk_status_binary ~ age, data = test_thera, family = binomial)
summary(model_age)$coef

model_sex <- glm(risk_status_binary ~ sex, data = test_thera, family = binomial)
summary(model_sex)$coef

model_surg <- glm(risk_status_binary ~ `type of surgery`, data = test_thera, family = binomial)
summary(model_surg)$coef

model_drain <- glm(risk_status_binary ~ `use of postoperative drain`, data = test_thera, family = binomial)
summary(model_drain)$coef

model_sinus <- glm(risk_status_binary ~ `entry of paranasal sinus`, data = test_thera, family = binomial)
summary(model_sinus)$coef

model_csf <- glm(risk_status_binary ~ `CSF leak`, data = test_thera, family = binomial)
summary(model_csf)$coef

model_duration <- glm(risk_status_binary ~ `duration of operation`, data = test_thera, family = binomial)
summary(model_duration)$coef

model_diabetes <- glm(risk_status_binary ~ `diabetes mellitus`, data = test_thera, family = binomial)
summary(model_diabetes)$coef

model_gcs <- glm(risk_status_binary ~ `GCS`, data = test_thera, family = binomial)
summary(model_gcs)$coef

model_ssi <- glm(risk_status_binary ~ `SSI`, data = test_thera, family = binomial)
summary(model_ssi)$coef

model_discharge <- glm(risk_status_binary ~ `discharge status`, data = test_thera, family = binomial)
summary(model_discharge)$coef

model_glucorticoids <- glm(risk_status_binary ~ `glucorticoids`, data = test_thera, family = binomial)
summary(model_glucorticoids)$coef

model_lumbar <- glm(risk_status_binary ~ `lumbar drainage`, data = test_thera, family = binomial)
summary(model_lumbar)$coef

model_income <- glm(risk_status_binary ~ `income bracket`, data = test_thera, family = binomial)
summary(model_income)$coef

model_bp <- glm(risk_status_binary ~ `blood pressure`, data = test_thera, family = binomial)
summary(model_bp)$coef



task11_model <- glm(risk_status_binary ~ age + sex + `CSF leak` +`duration of operation`+ GCS + `lumbar drainage` + `blood pressure`, data = test_thera, family = binomial)

summary(task11_model)$coef



task11_finalmodel <- glm(risk_status_binary ~ age + sex  +`duration of operation` + `lumbar drainage` + `blood pressure`, data = test_thera, family = binomial)

#Final Model
summary(task11_finalmodel)$coef

```


